---
title: "Fundamental Ratios"
author: "Atisha Rajpurohit"
date: '2022-04-05'
output: html_document
---

```{r}
library(tidyverse)
library(RColorBrewer)
```

```{r}
#setwd("~/Desktop/503/Project")
fundamental=read_csv('data/fundamentals.csv')
fcol=colnames(fundamental)
```

# Computing the Financial Ratios for 5 categories

## Liquidity Ratios

### 1. Current Ratio

This ratio is used to assess ther the company will be able to its bill over the next few months.
1 - Low
2 - Normal
3 - High

$$Current Ratio=Current Assets / Current Liabilities$$
columns - Total Current Assets/Total Current Liabilities

```{r}
fundamental2=fundamental %>% mutate(current_ratio=`Total Assets`/`Total Liabilities`) %>% 
  mutate(cr_int=case_when(current_ratio<1~"Low",
                          current_ratio==1~"Normal",
                          current_ratio>1~"High"))

```

### 2. Cash Ratio

This ratio assesses whether a company has enough cash to pay its current liabilties.
<1 Low
=1 Enough
>1 High

$$Cash Ratio=Cash and Cash Equivalents/Total Current Liabilties$$
column Cash Ratio

```{r}
fundamental2=fundamental2 %>% mutate(cash_ratio=`Cash Ratio`) %>% 
  mutate(ca_int=case_when(cash_ratio<1~"Low",
                          cash_ratio==1~"Normal",
                          cash_ratio>1~"High"))
```


$$ Quick Ratio$$

column Quick Ratio


## Financial Leverage

### 3. Asset Gearing

$$Leverage = Debt/Equity$$
columns - Long-Term Debt,Short-Term Debt / Current Portion of Long-Term Debt

```{r}
fundamental2=fundamental2 %>% mutate(leverage=(`Long-Term Debt`+`Short-Term Debt / Current Portion of Long-Term Debt`)/`Total Equity`) %>% 
  mutate(lev_int=case_when(leverage>0.5~"High",
                   ((leverage<=0.5)&(leverage>0.25))~"Normal",
                   leverage<=0.25~"Low"))
```


## Efficiency Ratios

### 4. Inventory Turnover Period

This ratio shows much inventory a company holds with respect to its cost of sales.

Interpretation
<2 - Low
2-4 - Normal
>4 - High

$$Inventory Turnover Ratio= Cost of Goods/Average Inventory$$
columns - Inventory/Cost of Revenue*365

```{r}
fundamental2=fundamental2 %>% mutate(inventory_turnover=(Inventory/`Cost of Revenue`)*365) %>% 
  mutate(inv_int=case_when(inventory_turnover<=20~"Low",
                           (inventory_turnover>20)|(inventory_turnover<40)~"Normal",
                           inventory_turnover>=40~"High"))
```



### 5. Asset Turnover Ratio

It tells a company how much revenue a company has generated based on the current assets.

Interpretation
<0.7 - Low
0.7-0.8 - Normal
>0.8 High

$$Asset Turnover Ratio = Total Revenue/Total Assets$$
```{r}
fundamental2=fundamental2 %>% mutate(asset_utilisation=`Total Revenue`/`Total Assets`) %>% 
  mutate(au_int=case_when(asset_utilisation<1~"Low",
                          asset_utilisation==1~"Normal",
                          asset_utilisation>1~"High"))
```


## Profitability Ratios

### 6. RoE


$$Return on Equity = Net Income/ Total equity$$

columns-Net Income, Total Equity

```{r}
fundamental2=fundamental2 %>% mutate(return_on_equity=`Net Income`/`Total Equity`) %>% 
mutate(roe_int=case_when(return_on_equity<0.1~"Low",
                         ((return_on_equity>=0.1)&(return_on_equity<0.15))~"Normal",
                         return_on_equity>=0.15~"High"))
```

### 7. RoA

$$Return on Assets = Net Income/ Total assets$$
```{r}
fundamental2=fundamental2 %>% mutate(return_on_Assets=`Net Income`/`Total Assets`) %>% 
mutate(roa_int=case_when(return_on_Assets<0.01~"Low",
                         ((return_on_Assets>=0.01)&(return_on_equity<0.05))~"Normal",
                         return_on_Assets>=0.05~"High"))
```


## Market Ratios

### 8. Price Earnings Ratio


$$Price Earnings Ratio = Market Price of a Share/Earnings per Share$$
For Market Price, the prices-adjusted-split dataset will be used, by joining on the dates.The closing prices will be used
```{r}
prices_adj=read_csv('data/prices-split-adjusted.csv')
#Merging the prices dataset with the fundamental dataset
fundamental3=fundamental2 %>% mutate(ticker_symbol=`Ticker Symbol`) %>% 
  mutate(period_ending=`Period Ending`)

market_prices=merge(fundamental3,prices_adj[,c("date","close","symbol")],by.x='ticker_symbol',by.y='symbol',all.x=TRUE)

fundamental3=market_prices %>% mutate(merger=case_when(period_ending==date~"True",
                                                  period_ending!=date~"False")) %>% 
  filter(merger=="True") %>% 
  mutate(price_earnings_ratio=close/`Earnings Per Share`) %>% 
  mutate(pe_int=case_when(price_earnings_ratio<15~"Low",
                          ((price_earnings_ratio>=15)&(price_earnings_ratio<18))~"Normal",
                          price_earnings_ratio>=18~"High"))
```


$$Dividend Payout Ratio= Dividend per Share/ Earning per Share$$
To calculate dividend payour ratio, the dividend 

```{r}
fundamental4=fundamental3 %>% select(cr_int,
                                     ca_int,
                                     lev_int,
                                     inv_int,
                                     au_int,
                                     roe_int,
                                     roa_int,
                                     pe_int)
fundamental5=fundamental3 %>% select(pe_int,
                                     lev_int)

fundamental6=fundamental3 %>% select(cr_int,
                                     ca_int,
                                     lev_int,
                                     inv_int,
                                     au_int,
                                     roe_int,
                                     roa_int,
                                     pe_int,
                                     ticker_symbol)

#write_csv(fundamental4,'ratio_test2.csv')
#write_csv(fundamental6,'ratio_test_radar.csv')

#write_csv(na.omit(fundamental3),'ratio_test_num.csv')
```

Creating a data frame for the circular barplot
```{r}
fundamental_4=fundamental3 %>% select(current_ratio,
                                     leverage,
                                     asset_utilisation,
                                     `After Tax ROE`,
                                     price_earnings_ratio,
                                     `For Year`,
                                     ticker_symbol)
write_csv(na.omit(fundamental_4),'5_ratios.csv')
```

```{r}
fundamental_5=fundamental3 %>% select(current_ratio,
                                     leverage,
                                     asset_utilisation,
                                     `After Tax ROE`,
                                     price_earnings_ratio,
                                     `For Year`,
                                     ticker_symbol,
                                     `Profit Margin`
                                     ) %>% 
  filter(`For Year`==2015)
#write_csv(na.omit(fundamental_4),'scatter_ratios.csv')
```